---
title: 'Shiny:: select input controlled by previous and next button'
author: Salekin
date: '2017-10-26'
slug: shiny-select-input-controlled-by-previous-and-next-button
categories:
  - visualization
tags:
  - Shiny
---

If you ever need to provide a select input in shiny that must be controlled by a previous and a forward button, you may consider using the following piece of code. A possible use case is to allow user to browse through different product with having to click on the dropdown menu and select the desired item. For a long list, this manual process can be tedious and a previous/forward button pair would make more sense.

### ui.R

```{r eval = FALSE}
# inspired by: https://gist.github.com/haozhu233/9dd15e7ba973de82f124
library(shiny)
library(shinyjs)
shinyUI(
  navbarPage("Demo", 
             position = "static-top",
             fluid = F,
             
             tabPanel("Demo",class="active",
                      sidebarLayout(
                        
                        sidebarPanel(
                          shinyjs::useShinyjs(),
                          
                          uiOutput("choose.date"),
                                     tags$div(class="row",
                                              tags$div(class="col-xs-6 text-center", uiOutput("prev_button.btn")),
                                              tags$div(class="col-xs-6 text-center", uiOutput("next_button.btn")))
                        ),
                        mainPanel = (
                          textOutput("dates.input")
                        )
                        )
                      )
             )
  )
```

### server.R


```{r eval = FALSE}
library(shiny)
library(shinyjs)

# inspired by: https://gist.github.com/haozhu233/9dd15e7ba973de82f124

shinyServer(function(input, output, session) {
  session$onSessionEnded(function() {
    stopApp()
  })
  
  choiceList <- c("Item 1", "Item 2", "Item 3")
  
  
  # ------- Select Input + previous/next week buttons---------------
  output$choose.date <- renderUI({
    selectInput("skus",
                "Select SKU",
                choices = choiceList)
  }) 
  
  output$prev_button.btn <- renderUI({
    actionButton("prev_button", 
                 label = HTML("<span class='small'><i class='glyphicon glyphicon-arrow-left'></i> Back</span>"))
  })
  output$next_button.btn <- renderUI({
    actionButton("next_button", 
                 label = HTML("<span class='small'>Next <i class='glyphicon glyphicon-arrow-right'></i></span>"))
  })
  

  observeEvent(input$prev_button, {
    if(input$skus == choiceList[1]){
      shinyjs::disable("prev_button")
      shinyjs::enable("next_button")
      updateSelectInput(session, "skus",
                        label = "Select SKU",
                        choices = choiceList,
                        selected = choiceList[1])
    } else{
      shinyjs::enable("prev_button")
      shinyjs::enable("next_button")
      updateSelectInput(session, "skus",
                        label = "Select SKU",
                        choices = choiceList,
                        selected = choiceList[match(input$skus, choiceList)-1])
    } 
    
    
    
    
    })
  
  observeEvent(input$next_button, {
    
    if(input$skus == choiceList[length(choiceList)]){
      shinyjs::disable("next_button")
      shinyjs::enable("prev_button")
      updateSelectInput(session, "skus",
                        label = "Select SKU",
                        choices = choiceList,
                        selected = choiceList[length(choiceList)])
    } else{
      shinyjs::enable("next_button")
      shinyjs::enable("prev_button")
      updateSelectInput(session, "skus",
                        label = "Select SKU",
                        choices = choiceList,
                        selected = choiceList[match(input$skus, choiceList)+1])
    } 
    
    
    
    })
  
  output$dates.input <- renderPrint({input$skus})
})


```
